\large Autonomous Exploration by Expected Information Gain from Probabilistic Occupancy Grid Mapping

### 2018-06-09 Conference/SIMPAR 2016/SIMPAR2016.tex
added Conference folder

Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. The popular occupancy grid mapping representation provides information about the free and occupied space as the robot measures its surroundings~\cite{ThrBurFox05}. Occupancy grid mapping is commonly used while estimating robot location and orientation, known as simultaneous localization and mapping (SLAM). Several SLAM approaches are applied to a variety of problems when mobile robots traverse through an environment to generate a map, but the trajectories are assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}). A major problem is that an effective trajectory would be unavailable knowing little or nothing about an environment. 
Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
In this paper, robotic motion is chosen to increase knowledge about the map; a robot chooses a trajectory designed to maximize information gain of the occupancy grid map.

@@

Approaches aimed at determining the future uncertainty of an occupancy grid map commonly suffer from several approximations.
%Determining future occupancy grid map uncertainty requires predictive estimations, which suffer from several approximations. 
Most evident is that all approaches rely on inaccurate probabilistic occupancy grids. This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}. Map-based information gain is commonly determined with Shannon's entropy~\cite{CarDamKumCas15}, an uncertainty metric based on grid cell occupancy probability. Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
Furthermore, the expected entropy of an action is not directly calculated~\cite{CarDamKumCas15,StaGriBur05}. Instead, these particle filter-based approaches assume that expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}. This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy. %Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.

In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}. This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
Thus, we provide the tools to approximate the exact solution that reduces the computational requirements with realistic assumptions.
The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
Compared with the recent results (e.g. [12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate solution to map information gain, under the assumption that the robot pose is known. In short, this paper presents an accurate and computationally-efficient approach to predict
map information gain for autonomous exploration with a novel structure that avoids common assumptions of the inverse sensor model and approximations in the entropy change.

The paper is organized as follows. The problem is formulated in Section II. The exact solution to the expected uncertainty due to an individual 1D measurement is derived in Section III, along with approximations to the solution. The concept of entropy is extended to 2D autonomous exploration with a numerical example in Section IV, followed by conclusions.

@@

The goal of occupancy grid mapping is to obtain the \emph{inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ for all grid cells inside the sensor field of view.
Suppose that the inverse sensor model before $t$-th time step is given, namely $P(m|X_{1:t-1},Z_{1:t-1})$, and the $l$-th measurement ray $z_{t,l}$ is available from pose $X_t$.
The inverse sensor model due to $z_{t,l}$ at the $t$-th time step follows Bayes' rule,
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})}\nonumber
\\
&=\eta_{t,l}p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $\eta_{t,l}$ is a normalizing constant independent of $m$, and we have used the fact that $X_t$ carries no information about $m$ without $z_{t,l}$. The issue is that $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ is perceived as intractable to calculate because the exact solution considers every combination of map $m$ to obtain the normalizer $\eta_{t,l}$, and therefore this approach yields $\mathcal O(2^{n_m})$ operations per grid cell. This perception is why heuristic solutions based on learning or intuition substitute the inverse sensor model. Recently, \refeqn{BayesRuleRayISM} was solved in a computationally-efficient manner to obtain an exact solution in real-time~\cite{KauLeeAiMos16}, which is summarized next.

@@

Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{CarDamKumCas15,StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
\begin{align}
\label{eqn:ShannonsEntropyDef}
H(&P(m))=\nonumber\\&-\sum_{i=1}^{n_m}\braces{
P(\mathbf{m}_i)\log{P(\mathbf{m}_i})+P(\bar{\mathbf{m}}_i)\log{(P(\bar{\mathbf{m}}_i}))}.
\end{align}
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized when $P(\mathbf{m}_i)\in\braces{0,1}$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

@@

The goal of autonomous exploration is to determine future poses that maximize the map information gain, or equivalently minimize the map entropy. 
%Let the objective function from the first time step $t=1$ to the $t_f$-th time step be
%\begin{align}
%\label{eqn:ObjFun}
%\mathcal I(X_{2:t_f})&=H(P(m|X_{1},Z_{1}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t_f},Z_{1:t_f}))\right].
%\end{align}
Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before the robot takes measurements of the entire reachable space.
Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.

Suppose at time $t$, $P(m|X_{1:t},Z_{1:t})$ is given. Let $X_c$ be the pose of the robot after the $t$-th time step.
The expected information gain $\mathcal I(X_c)$ is defined as the negative change
\begin{align}
\label{eqn:ObjFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right],
\end{align}
where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and $\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy when the robot is located at $X_c$, where the expectation is taken over $Z_c$.
Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
\begin{align}
\label{eqn:Objective}
X_c^*=\argmax_{X_c}{\mathcal I(X_c)},
\end{align}
where $X_c^*$ satisfies the inequality constraint to avoid collisions,
\begin{align}
\label{eqn:CollisionInequalityConstraint}
P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}(P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta,
\end{align}
where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a small acceptable probability of collision.
Once the robot has translated to $X_c^*$, the above process is repeated.
%Once the trajectory satisfying \refeqn{CollisionInequalityConstraint} between $X_t$ and $X_c$ is completed, the greedy process is repeated with an updated map.

@@

In this section, we present a method to determine the expected entropy from an individual measurement ray from a known pose. This method is based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.

\subsection{Single Ray Expected Value of Entropy}
We consider the expected entropy due to the $l$-th ray $z_{c,l}$ at position $x_c$,
\begin{align}
\label{eqn:HRayInt}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&
=\int_{z_\text{min}}^{z_\text{max}}
H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))\nonumber\\&\qquad\times p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
dz_{c,l}.
\end{align}
%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges, similar with how occupancy grid mapping decomposes continuous space into grid cells of length $\alpha$.
The discretized expected value of \refeqn{HRayInt} is
\begin{align}
\label{eqn:DiscExpEntropyRay}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\nonumber\\&\qquad\times P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
\end{align}
where index $z_{c,l,k}$ is the distance from $x_c$ to the $k$-th grid cell of the $l$-th reduced map $r_l$, known from geometry.

Standing alone, the term $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ from \refeqn{DiscExpEntropyRay} has a convoluted meaning because the depth $z_{c,l,k}$ does not directly depend on the map. However, we present a method to obtain this discretized probability. %Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
Following the assumption that $z_{c,l}$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
\begin{align}
\label{eqn:ProbWithDelta}
P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)}.
\end{align}
Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM},
\begin{align}
\label{eqn:DiscretizedProb}
p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\eta_{c,l,k}^{-1},
\end{align}
where $\eta_{c,l,k}^{-1}$ is defined in \refeqn{allEta}. However, instead of evaluating the $t$-th pose $X_t$ and measured ray $z_{t,l}$, we simply consider $X_c$ and ray depth $z_{c,l,k}$ based on histories $X_{1:t}$ and $Z_{1:t}$ to obtain the inverse normalizer,
\begin{align}
\label{eqn:allEtaAtC}
\eta_{c,l,k}^{-1}
&=
\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}\nonumber\\&\qquad\times p(z_{c,l,k}|\mathbf{r}_{l,i+},X_c)P(\mathbf{r}_{l,i}|X_{1:t},Z_{1:t}).
\end{align}
By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRay}, the expected entropy at $X_c$ from individual measurement ray $z_{c,l}$ is
\begin{align}
\label{eqn:HRayComplete}
&\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]=\left(\sum_{i=1}^{n_{r,l}+1}\eta_{c,l,i}^{-1}\right)^{-1}\nonumber\\&\quad\quad\times\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\eta_{c,l,k}^{-1}\bigg\},
\end{align}
where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
Here, \refeqn{HRayComplete} provides the expected entropies for those cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
The entropies of cells outside the field of view remain unchanged.

@@

The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}. Such computational requirements might be too cumbersome for some processors, especially if the measurement ray is expected to intersect many grid cells. Many of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.

The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially. This goal is achieved by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
The smaller set is determined by the probability that each cell is captured by the measurement ray, known as the detection probability. This can be found recursively as
%These cells correspond to the $\hat n$ cells with the greatest probabilities of being captured by the measurement ray, known as the detection probability. This can be found recursively as

@@

Let $\hat{r}_{l}$ correspond to the grid cells that yield the $\hat{n}$ maximum values \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.

@@

Even though the value of $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, so the computational order is fixed as well.

@@

@@

We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}). The variable $a_\text{temp}$ is an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification. It can be noted that if $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.

@@

The purpose of this numerical example is to provide evidence that the approximations are reasonable and speed up the algorithm substantially.
Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have spacing $\alpha=0.2$ meters, and the properties of the range sensor are based off the Microsoft Kinect~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV). The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, which only considers $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.

@@

The robot pose at a future time step is selected to maximize \refeqn{ObjFun} over the search space $\Re^2\times\Sph^1$. We reduce the search by discretizing the search space in both position and attitude as follows. First, we choose $n_c$ candidate locations evenly-spaced along a circle of radius $\delta$, centered around the current location $x_t$, where the $c$-th candidate location is denoted $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$, illustrated in Figure \ref{fig:OptProcess}. Any location that violates \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions. For each collision-free candidate location, we consider $n_d$ evenly-spaced directions. The rays and attitudes at $x_c$ are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$ and $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, respectively, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV). For each candidate location, we choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
\begin{align}
\label{eqn:FindRc}
&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
\end{align}

>>>>

Robotic mapping is the process of generating maps, which represent the environment in close proximity of a robot. The popular occupancy grid mapping representation provides information about the free and occupied space as the robot measures its surroundings~\cite{ThrBurFox05}. Occupancy grid mapping is commonly used while estimating robot location and orientation, known as simultaneous localization and mapping (SLAM). Several SLAM approaches are applied to a variety of problems when mobile robots traverse an environment to generate a map, but the robot actions assumed given (e.g.~\cite{ThrBurFox05,DurBai06,CheChe09}). A major problem is that generating an effective trajectory for the robot to follow would be unavailable knowing little or nothing about an environment. 
Autonomous exploration solves this problem, by choosing robotic motion through uncertain space, while building a map that is used for subsequent navigation~\cite{Yam97}.
In this paper, robotic motion is chosen to increase knowledge about the map; a robot chooses a trajectory designed to maximize the map information gain. % of the occupancy grid map.

@@

Approaches aimed at determining the future uncertainty of an occupancy grid map commonly suffer from several approximations.
%Determining future occupancy grid map uncertainty requires predictive estimations, which suffer from several approximations. 
Most evident is that all approaches rely on inaccurate probabilistic occupancy grids. This mapping representation is typically based on heuristic approximations (e.g.~\cite{MorElf85,Elf89,ChoLynHutKanBurKavThr05,And09,PirRutBisSch11}) or simulated learning techniques \cite{ThrBurFox05,Thr01}. Map-based information gain is commonly determined with Shannon's entropy~\cite{StaGriBur05}, an uncertainty metric based on grid cell occupancy probability. Since this probability is heuristic or learned, any measure of entropy is subject to errors from approximated probability.
Furthermore, the expected entropy of an action is not directly calculated. Instead, these particle filter-based approaches assume expected entropy is equivalent to entropy based on expected measurement scans~\cite{JohStaPfaBur07}. This assumption is subject to inaccuracies due to the nonlinear nature of Shannon's entropy. %Furthermore, these approaches typically generate maps with a particle filter, which commonly require some tuning and may consume large amounts of onboard memory.

In this paper, we propose an autonomous exploration technique based on expected entropy change that avoids the aforementioned issues. 
%The occupancy grid mapping scheme is based on~\cite{KauLeeAiMos16}, which provides the exact solution of occupancy grid mapping of a single measurement ray. 
We develop a solution for expected entropy gain based on exact occupancy grid mapping~\cite{KauLeeAiMos16}. This solution directly calculates expected entropy from a probabilistic map based on similar assumptions of those applied to occupancy grid mapping.
Though more accurate than several other approaches, the proposed solution requires large computational resources, making real-time implementation difficult in certain scenarios.
Thus, we provide the tools to approximate the exact solution that reduces the computational requirements with realistic assumptions.
The motion of the robot is formulated as an optimization problem, where the expected entropy change is minimized, or equivalently the expected information gain is maximized.
Compared with the recent results (e.g. [12], [13], [16], [17]) that employ Rao-Blackwellized particle filters to deal with localization and autonomous exploration simultaneously with an approximate entropy change, this paper is based on an accurate solution to map information gain, under the assumption that the robot pose is known. In short, this paper presents an accurate and computationally-efficient approach to predict
map information gain for autonomous exploration with a novel structure that avoids common assumptions of the inverse sensor model and approximations of entropy change.

The paper is organized as follows. The problem is formulated in Section II. The exact solution to the expected uncertainty due to an individual 1D measurement is derived in Section III, along with optional approximations. The concept of entropy is extended to 2D autonomous exploration with a numerical example in Section IV, followed by conclusions.

@@

The goal of occupancy grid mapping is to obtain the \emph{inverse sensor model} $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$ for all grid cells inside the sensor field of view.
Suppose that the inverse sensor model before $t$-th time step is given, namely $P(m|X_{1:t-1},Z_{1:t-1})$, and the $l$-th measurement ray $z_{t,l}$ is available from pose $X_t$.
The inverse sensor model due to $z_{t,l}$ at the $t$-th time step follows Bayes' rule,
\begin{align}
\label{eqn:BayesRuleRayISM}
P(&m|z_{t,l},X_{1:t},Z_{1:t-1})\nonumber
\\
&=\frac{p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1})}{p(z_{t,l}|X_{1:t},Z_{1:t-1})}\nonumber
\\
&=\eta_{t,l}p(z_{t,l}|m,X_{1:t},Z_{1:t-1})P(m|X_{1:t-1},Z_{1:t-1}),
\end{align}
where $\eta_{t,l}$ is a normalizing constant independent of $m$, and we used the fact that $X_t$ carries no information about $m$ without $z_{t,l}$. However, $P(\mathbf{m}_i|z_{t,l},X_{1:t},Z_{1:t-1})$ is perceived as intractable because the exact solution considers every combination of map $m$ to obtain the normalizer $\eta_{t,l}$, yielding $\mathcal O(2^{n_m})$ per grid cell. This is why heuristic solutions based on learning or intuition substitute the inverse sensor model. Recently, \refeqn{BayesRuleRayISM} was solved computationally-efficiently to obtain an exact solution in real-time~\cite{KauLeeAiMos16}, summarized next.

@@

Since occupancy grid mapping provides a probabilistic representation of surrounding space, this mapping scheme holds probabilistic information about the uncertainty of the map. Shannon's entropy is commonly used as a measure of uncertainty, such as in~\cite{StaGriBur05}. Given the probabilities of each grid cell of map $m$, Shannon's entropy is defined as the sum of the entropy of each cell,
\begin{align}
\label{eqn:ShannonsEntropyDef}
H(&P(m))=\nonumber\\&-\sum_{i=1}^{n_m}\braces{
P(\mathbf{m}_i)\log{P(\mathbf{m}_i})+P(\bar{\mathbf{m}}_i)\log{(P(\bar{\mathbf{m}}_i}))}.
\end{align}
The entropy of the $i$-th grid cell is maximized when $P(\mathbf{m}_i)=0.5$ (greatest uncertainty) and minimized when $P(\mathbf{m}_i)\in\braces{0,1}$ (smallest uncertainty). %; thus Shannon's entropy is a measure of uncertainty.

@@

The goal of autonomous exploration is to determine future poses that maximize the map information gain, or equivalently minimize the map entropy. 
%Let the objective function from the first time step $t=1$ to the $t_f$-th time step be
%\begin{align}
%\label{eqn:ObjFun}
%\mathcal I(X_{2:t_f})&=H(P(m|X_{1},Z_{1}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t_f},Z_{1:t_f}))\right].
%\end{align}
Since autonomous exploration is conducted in uncertain environments, a complete trajectory cannot be known before a priori.
Instead, an objective function governs the motion between intermediate poses, building an occupancy grid map along the way.

Suppose at time $t$, $P(m|X_{1:t},Z_{1:t})$ is given. Let $X_c$ be the pose of the robot after the $t$-th time step.
The expected information gain $\mathcal I(X_c)$ is defined as the negative change
\begin{align}
\label{eqn:ObjFun}
\mathcal I(X_c)&=H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\quad-\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right],
\end{align}
where $H(P(m|X_{1:t},Z_{1:t}))$ may be computed with \refeqn{ShannonsEntropyDef} and $\text{E}\left[H(P(m|X_{1:t},Z_{1:t},X_c,Z_c))\right]$ is the expected entropy with robot pose $X_c$ and the expectation is taken over $Z_c$.
Since autonomous exploration is formulated as a trajectory optimization, the goal is to repeatedly choose optimal $X_c^*$ that maximizes the objective function
\begin{align}
\label{eqn:Objective}
X_c^*=\argmax_{X_c}{\mathcal I(X_c)},
\end{align}
avoiding collisions by satisfying the inequality constraint
\begin{align}
\label{eqn:CollisionInequalityConstraint}
P_\text{collision}(X)=1-\prod_{i\in\mathcal C_X}(P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})\leq\beta,
\end{align}
where $\mathcal C_X$ is the set of grid cells falling inside a volume of preselected size around $X$ that may cause collision and $\beta>0$ is a small acceptable probability of collision.
Once the robot has translated to $X_c^*$, the above process is repeated.
%Once the trajectory satisfying \refeqn{CollisionInequalityConstraint} between $X_t$ and $X_c$ is completed, the greedy process is repeated with an updated map.

@@

The expected entropy from an individual measurement ray from a known pose is presented, based on the exact solution to occupancy grid mapping from~\cite{KauLeeAiMos16}.

\subsection{Single Ray Expected Value of Entropy}
We consider the expected entropy due to the $l$-th measurement ray $z_{c,l}$ at position $x_c$,
\begin{align}
\label{eqn:HRayInt}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&
=\int_{z_\text{min}}^{z_\text{max}}
H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))\nonumber\\&\qquad\quad\times p(z_{c,l}|X_{1:t},Z_{1:t},x_c)
dz_{c,l}.
\end{align}
%where $z_\text{min}$ and $z_\text{max}$ are sensor limits. Considering the continuous space of possible measurements of $z_{c,l}$ would make \refeqn{HRayInt} intractable to calculate. Thus, w
We discretize the measurement ray space such that $z_{c,l}$ falls on points along the measurement ray intersecting with grid cell edges.
The discretized expected value of \refeqn{HRayInt} is
\begin{align}
\label{eqn:DiscExpEntropyRay}
\text{E}[H(P(m|&X_{1:t},Z_{1:t},x_c,z_{c,l}))]
\nonumber\\&=\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\nonumber\\&\qquad\qquad\times P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)\bigg\},
\end{align}
where index $z_{c,l,k}$ is the distance from $x_c$ to the $k$-th grid cell of the $l$-th reduced map $r_l$, known from geometry.

Standing alone, the term $P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)$ from \refeqn{DiscExpEntropyRay} has a convoluted meaning because the depth $z_{c,l,k}$ does not directly depend on the map. However, we present a method to obtain this discretized probability. %Despite how this term is replaced with a normalizer in ~\cite{ThrBurFox05} replaces
Following the assumption that $z_{c,l}$ is discretized to known distances, the probabilities are proportional to their densities, as the area under the density curve is infinitesimal and fixed. Accounting for all cases, the probability is
\begin{align}
\label{eqn:ProbWithDelta}
P(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)=\frac{p(z_{c,l,k}|X_{1:t},Z_{1:t},x_c)}{\sum_{i=1}^{n_{r,l}+1}p(z_{c,l,i}|X_{1:t},Z_{1:t},x_c)}.
\end{align}
Conveniently, this density corresponds to the inverse normalizer defined in \refeqn{BayesRuleRayISM} as the denominator from Bayes' rule,
\begin{align}
\label{eqn:DiscretizedProb}
p(z_{c,l,k}|X_{1:t},X_c,Z_{1:t})=\eta_{c,l,k}^{-1},
\end{align}
where $\eta_{c,l,k}^{-1}$ is defined in \refeqn{allEta}. However, instead of evaluating the $t$-th pose $X_t$ and measured ray $z_{t,l}$, we simply consider $X_c$ and ray depth $z_{c,l,k}$ based on histories $X_{1:t}$ and $Z_{1:t}$ to obtain the inverse normalizer,
\begin{align}
\label{eqn:allEtaAtC}
\eta_{c,l,k}^{-1}
&=
\sum_{i=1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t},Z_{1:t})\bigg\}\nonumber\\&\qquad\times p(z_{c,l,k}|\mathbf{r}_{l,i+},X_c)P(\mathbf{r}_{l,i}|X_{1:t},Z_{1:t}).
\end{align}
By substituting \refeqn{ProbWithDelta} and \refeqn{DiscretizedProb} into \refeqn{DiscExpEntropyRay}, the expected entropy at $X_c$ from individual measurement ray $z_{c,l}$ is
\begin{align}
\label{eqn:HRayComplete}
&\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]=\left(\sum_{i=1}^{n_{r,l}+1}\eta_{c,l,i}^{-1}\right)^{-1}\nonumber\\&\quad\quad\times\sum_{k=1}^{n_{r,l}+1}\bigg\{H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l,k}))\eta_{c,l,k}^{-1}\bigg\},
\end{align}
where $\eta_{c,l,k}^{-1}$ is taken from \refeqn{allEtaAtC}.
Here, \refeqn{HRayComplete} provides the expected entropies for cells inside the field of view of $r_l$. %; cells outside this reduced map are trivially changed by $0$ due to $z_{c,l}$.
The entropies of cells outside the field of view remain unchanged.

@@

The order of computation for each measurement ray is $\mathcal O((n_{r,l}+1)^2)$ since the summations of \refeqn{allEtaAtC} are embedded in \refeqn{HRayComplete}. However, several of those intersections provide negligible information since the probability of the $l$-th measurement ray capturing certain cell depths is close to zero.

The approximation of expected ray entropy provides a method to reduce the computation of \refeqn{allEtaAtC} substantially by systematically selecting a smaller set of grid cells to consider over the summations of \refeqn{allEtaAtC} and \refeqn{HRayComplete}.
The smaller set is determined by the detection probability, or the probability that a particular cell is captured by the measurement ray, %These cells correspond to the $\hat n$ cells with the greatest probabilities of being captured by the measurement ray, known as the detection probability. This can be found recursively as

@@

Let $\hat{r}_{l}$ correspond to grid cells that yield the $\hat{n}$ maximum values \refeqn{ProbOfFirstCell} (the $\hat n$ most likely ray detections), indexed by increasing distance from candidate location $x_c$.

@@

Even though $n_{r,l}$ is different among various rays in general, $\hat n$ is fixed, which fixes the computational order.

@@

%When $\hat n=n_{r,l}+1$, the ray objective function is computed without approximation.

@@

We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Table \ref{tab:RayExpectedEntropyGain}). The variable $a_\text{temp}$ is an intermediate variable designed to avoid repeated calculations.
Fixed indices and condition variables are removed for simplification.

@@

The purpose of this numerical example is to provide evidence that the approximations are reasonable and reduce computation substantially.
Since a measurement ray produces a depth measurement in a single direction, we only consider a 1D map where the grid cells have edge length $\alpha=0.2$ meters. We apply properties of the Microsoft Kinect depth sensor~\cite{PirRutBisSch11,KhoElb12} with maximum reading depth $z_\text{max}=4$m ($20$ grid cells inside the sensor FOV). The goal is to compare the expected entropy $\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$ from \refeqn{HRayComplete} and with an approximation $\text{E}[H_\text{approx}(P(m|X_{1:t},Z_{1:t},x_c,z_{c,l}))]$, only considering $\hat n$ grid cells with highest detection probability \refeqn{ProbOfFirstCell}.

@@

The robot pose at a future time step is selected to maximize \refeqn{ObjFun} over the search space $\Re^2\times\Sph^1$. We reduce the search by discretizing the search space in both position and attitude as follows. First, we choose $n_c$ candidate locations evenly-spaced along a circle of radius $\delta$, centered around the current location $x_t$, where the $c$-th candidate location is denoted $x_c\in\braces{x_1,x_2,\ldots,x_{n_c}}$, illustrated in Figure \ref{fig:OptProcess}. Any location that violates \refeqn{CollisionInequalityConstraint} is excluded to avoid collisions. For each candidate location, we consider $n_d$ evenly-spaced directions. The rays and attitudes at $x_c$ are denoted $\braces{z_{c,1},z_{c,2},\ldots,z_{c,n_d}}$ and $\braces{R_{c,1},R_{c,2},\ldots,R_{c,n_d}}$, respectively, where the scan with attitude $R_{c,d}$ might cover several ray directions depending on the sensor field of view (FOV). At $x_c$, we choose optimal attitude $R_c^*$ as the summation of the expected entropy changes covered by the scan,
\begin{align}
\label{eqn:FindRc}
&R_c^*=\argmax_{R_{c,d}}\sum_{z_{c,i}\in R_{c,d}\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\qquad\quad\quad\quad\quad-\text{E}[H(P(m|X_{1:t},Z_{1:t},x_c,z_{c,i}))]\bigg).
%\mathcal I_\text{ray}(x_c,z_{c,i})\right).
\end{align}

Therefore, the optimal attitude $R_c^*$ is given as a function of the location $x_c$. The objective function \refeqn{ObjFun} is computed by a summation about the $n_d$ rays as %\refeqn{Objective},
\begin{align}
\label{eqn:ObjFunApprox}
\mathcal I(x_c,R_c^*)&\approx \sum_{z_{c,i}\in R_{c}^*\text{ FOV}}\bigg(H(P(m|X_{1:t},Z_{1:t}))\nonumber\\&\ -\text{E}\left[H(P(m|X_{1:t},Z_{1:t},x_c,R_c^*,z_{c,i}))\right]\bigg),
\\
x_c^*&=\argmax_{x_c}\mathcal I(x_c,R_c^*).
\end{align}
%such that $X_c^*=\braces{x_c^*,R_c^*}$ is the optimal candidate over the discretized search space. Figure \ref{fig:OptProcess} illustrates this process.
